\documentclass[12pt]{article}
\usepackage{fullpage}
\usepackage{titlesec}
\usepackage{tikz}
\usepackage{amsfonts,amssymb}
\usepackage{amsmath}
\usepackage{comment}
\relpenalty=9999
\binoppenalty=9999

\begin{document}
\pagestyle{plain}
\titleformat{\subsection}[runin]
  {\normalfont\large\bfseries}{\thesubsection}{1em}{}
\titleformat{\subsubsection}[runin]
  {\normalfont\large\bfseries}{\thesubsubsection}{1em}{}

\section*{Problem 3a.}
Without loss of generality let each clause in $\Phi$ have exactly $3$
distinct literals. Let $S$ be a set of $C$ disjoint clauses in $\Phi$.
A single clause in $\Phi$ (and $S$ respectively) can be satisfied by $7$ of the
$8$ possible assingments (since there is only one assignment that does
not satisfy the clause i.e. the one where each of the $3$ literals evaluates to
false). Since the clauses in $S$ are disjoint, an assignment for the variables
in one clause does not affect the satisfiabilty of another clause. Therefore,
there are exactly $7^C$ satisfying assignments out of the total $8^C$ possible
assignments for the set $S$. For $\Phi$ to be satisfied, both the variables in
$S$ and the variables not in $S$ need to have a satisfying assignment. Now
consider the remaining $n-3C$ variables not in $S$. At best, all the $2^{n-3C}$
possible assignments satisfy the clauses not in $S$. Combining all the possible
satisfying assignments for the two sets there are at most $7^C \cdot 2^{n-3C}$
possible satisfying assignments for $\Phi$ out of the total $2^n$ possible
assignments. To satisfy the condition that at least $1\%$ of the possible
assignments satisfy all the clauses in the formula $\Phi$ we have the following
inequality:
$$\frac{7^C \cdot 2^{n-3C}}{2^n} \ge \frac{1}{100} $$
which gives a threshold of $34$, after which the inequality does not hold
and the fact that $1\%$ of the assignments satify $\Phi$ is contradicted.

\section*{Problem 3b.}
Consider the following algorithm for finding a set $V$ of at most $3C$ variables
with the property that every clause in $\Phi$ has at least one variable in $V$.
Maintain a list of the clauses each variable belongs to. Choose the variable that
belongs to the most number of clauses to be in $V$. Repeat this process, modulo
the clauses that are already covered by selected variables, until every clause
has at least one variable in $V$. This algorithm takes $O(m)$ preprocessing time
to create the list of variable-clause relationships. The algorithm runs in at
most $O(n)$ time (to remove the newly covered clauses from the variable-clause
relationship list) since the algorithm iterates at most $3C$ times (which is yet
to be shown). Finally we can show that the size of $V$ is at most $3C$. Consider
the largest set $S$ of disjoint clauses in $\Phi$. From part (a) we know that
$|S| \le C$ therefore $S$ contains at most $3C$ variables. Since $S$ is the
largest set of disoint clauses, every clause in $\Phi$ has at least one variable
in $S$. Proof by contradiction: if there exists a clause with no variables in
$S$ then that clause can be added to $S$ since it is disjoint from every other
clause in $S$, which increases the size of $S$ to $|S|+1$ therefore the orignal
set $S$ was not the largest set of disjoint clauses, a contradiction. Now
consider a set $S' \subseteq S$ of all the variables in $S$ that cover every
clause in $\Phi$. Clearly each of these variables covers a number of clauses in
$\Phi$ such that all of $\Phi$ is covered by them. Since the above algorithm
choses variables that cover the maximum number of clauses to cover $\Phi$, the
algorithm would select, at most, all the variables in $S'$ which is of size at
most $3C$. Otherwise it would chose other variables in lieu of variables in $S'$
that cover more clauses, thereby needing less total variables to cover $\Phi$.
The correctness of the algorithm is straightforward since it runs until every
clause in $\Phi$ is covered by at least one variable in $V$.

\section*{Problem 3c.}
Using part (b) we can construct a set $V$ of at most $3C$ variables with the
property that every clause in $\Phi$ has one variable in $V$. Now consider
chosing one of the $2^{3C}$ possible assignments for the variables in $V$. We
can use this assignment to reduce $\Phi$ in the following way. First, we can
remove the clauses that are satisfied by the assignment. Second, note that all
of the remaining clauses (the ones that are still not satisfied) have at least
one variable in $\Phi$. Therefore, we can remove that variable from each
of the unsatisfied clauses (since it has an assignment and that assignment does
not satisfy the clause) and reduce those clauses to 2-CNF clauses. Now we
are left with an 2-SAT instance which we can solve in polynomial time. If that
algorithm finds a satisfying assignment then we can combine it with the original
assignment for the $3C$ variables in $V$ to get a satisfying assignment for
$\Phi$. Otherwise, we repeat this procedure, chosing a new assignment for the
variables in $V$. As noted above, this algorithm will iterate at most $2^{3C}$
times, once for each possible assignment for the variables in $V$, until the
algorithm returns a satisfying assignment. Therefore, the algorithm needs to
solve at most $2^{3C}$ 2-SAT instances. We can reduce $\Phi$ in $O(m)$ time
and solve a 2-SAT intance in polynomial time therefore the entire algorithm
runs in polynomial time including the preprocessing construction of the set
$V$ from part (b).


\end{document}
