\documentclass[12pt]{article}
\usepackage{fullpage}
\usepackage{titlesec}
\usepackage{tikz}
\usepackage{amsfonts,amssymb}
\usepackage{amsmath}
\usepackage{comment}
\relpenalty=9999
\binoppenalty=9999

\begin{document}
\pagestyle{plain}
\titleformat{\subsection}[runin]
  {\normalfont\large\bfseries}{\thesubsection}{1em}{}
\titleformat{\subsubsection}[runin]
  {\normalfont\large\bfseries}{\thesubsubsection}{1em}{}

\section*{Problem 3.}
Consider a satisfying 3-coloring assignment $A$ of the graph $G$ using the
colors red, blue, and green. Any triangle must have each of its vertices
colored differently. Now consider a relaxed 2-coloring assignment $A'$ where
every green vertex in $A$ is arbitrarily colored either red or blue. $A'$
still satisfies the constraint that no triangle is monocromatic. The proof of
this is straightforward. Every triangle contains a red, blue, and green vertice
in $A$. Changing all the green vertices in $A$ to be either red or blue means
that there is still at least one red and at least one blue vertice in every
triangle since the original red and blue vertices remained unchanged from $A$ to
$A'$. Fix such a coloring $A'$ and consider the following algorithm.
Arbitrarily assign a 2-coloring to the graph $G$. While there remain
mono-cromatic triangles in $G$, arbitrarily select one, randomly choose one of
its vertices and flip its color. We repeat until every vertice in a triangle has
the color assigned to it in $A'$ (or until we satisfy the constraint, which we
know at least happens when our coloring matches $A'$). We know from above that
each triangle in $G$ has one vertice who's color must be blue, one vertice who's
color must be red and one vertice who's color does not matter (the vertice who's
color would have been green in the 3-coloring $A$ of the graph). Therefore, the
probability that we choose a correctly colored vertice and flip its color is
equal to the probability that we choose an incorrectly colored vertice and flip
its color which is equal to the probability we choose a vertice who's coloring does
not affect the constraint satisfaction which is equal to $1/3$. We can model
this algorithm as the random walk of a path graph defined in the following way:
each vertex $v_i$ has an edge to vertex $v_{i-1}$, $v_{i+1}$, and a self loop
each with an associated probability of $1/3$. Being at a vertice $v_i$ corresponds
to $i$ vertices in $G$ being correctly colored according to our satisfying assignment $A'$.
We have a path of length at most $n$ since we only need to consider the vertices who are
members of a triangle, which is at most all the vertices in $G$. The cover time
of our path graph is at most $2\cdot(3n)\cdot(n-1) = 6n^2+6n$ steps, therefore
our algorithm takes $O(n^2)$ expected running time.

\end{document}
